---
title: "Laboratorio 01"
output:
  pdf_document: default
date: "2023-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bloque 0

### Problema 1

```{r, p1}
  x = c(1,2,3,4)
  y = seq(1:10)
  z = seq(2,10,2)
  w = rep(0,10)
  y[3] = 1
```

### Problema 2
```{r, p2}
  curve(x+1, from=1, to=10 ,xlab="x", ylab="y")
  curve((x^2)+1, from=1, to=10, xlab="x", ylab="y")
  curve(sin(x), from=1, to=10, xlab="x", ylab="y")
  curve(cos(x), from=1, to=10, xlab="x", ylab="y")
  curve(log(x), from=1, to=10, xlab="x", ylab="y")
```


### Problema 3
```{r,p3}
  curve(x+1, from=1, to=10 ,xlab="x", ylab="y")
  curve((x^2)+1, from=1, to=10, xlab="x", ylab="y")
  curve(sin(x), from=1, to=10, xlab="x", ylab="y")
  curve(cos(x), from=1, to=10, xlab="x", ylab="y")
  curve(log(x), from=1, to=10, xlab="x", ylab="y")
```


### Problema 4
```{r, p4}
library(readr)
dat <- read_csv("./pizzadelivery.csv")
  #str(dat)
  #dim(dat)
  #colnames(dat)
  #names(dat)
  #nrow(dat)
  #ncol(dat)
  #head(dat)
  #tail(dat)
  dat["temperature"] = (dat["temperature"] * (9/5)) + 32
  dat[dat$driver == 'Domenico',]
  dat[dat$pizzas >= 4,"day"]
```

## Bloque 1

### Problema 5
- day : cualitativa-discreta
- date : cualitativa-ordinal
- time : cuantitativa-discreta
- operator : cualitativa-nominal
- branch : cualitativa - ? 
- driver : cualitativa-nominal
- temperature : cuantitativa-continua
- bill : cuantitativa-discreta
- pizzas : cuantitativa-discreta
- free_wine : cualitativa-nominal
- got_wine : cualitativa-nominal
- discount_customer : cualitativa- ?

### Problema 6

```{r p6}
  fi <- table(dat$driver)
  hi <- fi/length(dat$driver)
  barplot(fi)
  tabla <- cbind(fi,hi)
```
### Problema 7

```{r}
table(dat[dat$free_wine==TRUE,7])
```

### Problema 8

```{r}
table(dat[dat$free_wine==T & dat$got_wine==T,6])
```


## Bloque 2

### Problema 1

```{r}
    # temperature
    library(readr)
    db <- read_csv("./pizzadelivery.csv")
    # Variable: Time
    temp_list <- db$temperature;
    
    # Calculate information to create frequency table
    min<- min(temp_list);
    max <- max(temp_list);
    length <- length(temp_list);
    range <- max - min;
    class_num <- round(1 + log(length, 2));
    class_width <- round(range / class_num, 2);
    
    print(c("Min", min));
    print(c("Max", max));
    print(c("Length", length));
    print(c("Range", range));
    print(c("Class Number", class_num));
    print(c("Class Width", class_width));
    
    #Calculate intervals and midpoints
    sequ <- seq(min, max + class_width, class_width);
    intervals <- cut(temp_list, sequ, right = TRUE, include.lowest = TRUE);
    abs_freq <- table(intervals);
    midpoints <- round((head(sequ, -1) + diff(sequ) / 2), 2);
    
    #Generate frequency table
    freq_table = cbind(
        Mark = midpoints,
        Abs = abs_freq, 
        Reltv = round(abs_freq / length(temp_list), 2), 
        Acm = cumsum(abs_freq),
        ReAcm = round(cumsum(abs_freq) / length(temp_list))
    );
    
    print(freq_table);
    #fi = hist(dat$temperature)
```


```{r}
  #time
  min = min(dat$time)
  max = max(dat$time)
  k = 5
  A = (max-min)/ k
  
  intervals = cut(dat$time, breaks=seq(min, max, by = A), Right = TRUE, include.lowest=TRUE)
  fi = table(intervals)
  
  porc = (fi/ length(dat$time))*100 
  cbind(fi,porc)
  
```


```{r}
  # bill
  fi = hist(dat$bill)
```

```{r}
  fi = table(dat$pizzas)
  hi  = fi/length(dat$pizzas)
  # porc = (fi/ length(dat$pizzas))*100 
  barplot(fi)
  
```
