---
title: "Laboratorio 01"
Authors: "Katherine Benjumea Ortiz, Brigith Lorena Giraldo Vargas, Valentina Ochoa Arboleda, Tomás Atehortua Ceferino, Danilo de Jesús Toro Echeverri" 
output:
  html_document: default
  pdf_document: default
date: "2023-02-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bloque 0

### Problema 1

```{r, p1}
  x = c(1,2,3,4)
  y = seq(1:10)
  z = seq(2,10,2)
  w = rep(0,10)
  y[3] = 1
```

### Problema 2

```{r, p2}
  curve(x+1, from=1, to=10 ,xlab="x", ylab="y")
  curve((x^2)+1, from=1, to=10, xlab="x", ylab="y")
  curve(sin(x), from=1, to=10, xlab="x", ylab="y")
  curve(cos(x), from=1, to=10, xlab="x", ylab="y")
  curve(log(x), from=1, to=10, xlab="x", ylab="y")
```

### Problema 3

```{r,p3}
  curve(x+1, from=1, to=10 ,xlab="x", ylab="y")
  curve((x^2)+1, from=1, to=10, xlab="x", ylab="y")
  curve(sin(x), from=1, to=10, xlab="x", ylab="y")
  curve(cos(x), from=1, to=10, xlab="x", ylab="y")
  curve(log(x), from=1, to=10, xlab="x", ylab="y")
```

### Problema 4

```{r, p4}
library(readr)
dat <- read_csv("C:/Users/Peanut/Desktop/Estadistica/Laboratorio1/pizzadelivery.csv")
  #str(dat)
  #dim(dat)
  #colnames(dat)
  #names(dat)
  #nrow(dat)
  #ncol(dat)
  #head(dat)
  #tail(dat)
  dat["temperature"] = (dat["temperature"] * (9/5)) + 32
  dat[dat$driver == 'Domenico',]
  dat[dat$pizzas >= 4,"day"]
```

## Bloque 1

### Problema 5

- day : cualitativa-nominal
- date : cuantitativa-discreta
- time : cuantitativa-continua
- operator : cualitativa-nominal
- branch : cualitativa-nominal
- driver : cualitativa-nominal
- temperature : cuantitativa-continua
- bill : cuantitativa-continua
- pizzas : cuantitativa-discreta
- free_wine : cuantitativa-discreta
- got_wine : cuantitativa-discreta
- discount_customer : cuantitativa-discreta

### Problema 6

# - day : cualitativa-nominal
```{r p6}
(fi_day<-table(dat$day))
(porc_2<-(fi_day/length(dat$day))*100)
porc_2<-round(porc_2,2)
pie(porc_2, porc_2, col = palette("Set 2"))
legend(x ="bottomright", legend = names(fi_day), fill = palette("Set 2"))
```

#- operator : cualitativa-nominal
#preguntar etiqueta de barra
```{r}
(fi_operator<-table(dat$operator))
(porc_2<-(fi_operator/length(dat$operator))*100)
porc_2<-round(porc_2,2)
barplot(porc_2, porc_2, col=4:6)
legend(x ="topright", legend = names(fi_operator), fill = 4:6)
```

#- branch : cualitativa-nominal
```{r}
library(plotrix)
(fi_branch<-table(dat$branch))
(porc_2<-(fi_branch/length(dat$branch))*100)
porc_2<-round(porc_2,2)
pie3D(porc_2, labels = porc_2, col=2:9)
legend(x ="bottomleft", legend = names(fi_branch), fill = 2:9)
```

#- driver : cualitativa-nominal
#preguntar cambiar grosor etiqueta de barra
```{r}
(fi_driver<-table(dat$driver))
(porc_2<-(fi_driver/length(dat$driver))*100)
porc_2<-round(porc_2,2)
barplot(porc_2, porc_2, col=3:8, horiz = TRUE, las=TRUE)
legend(x ="topleft", legend = names(fi_driver), fill = 3:8)
```

### Problema 7

```{r}
library(plotrix)      
datp7<-dat[dat$free_wine==1,]
(p7<-table(datp7$driver))
(porc_p7<-(p7/length(datp7$driver))*100)
porc_p7<-round(porc_p7,2)
pie3D(porc_p7, labels = porc_p7, col=2:9)
legend(x ="top", legend = names(p7), fill = 2:9)
```

### Problema 8

```{r}
datp8<-dat[dat$free_wine==1&dat$got_wine==1,]
(p8<-table(datp8$branch))
(porc_p8<-(p8/length(datp8$branch))*100)
porc_p8<-round(porc_p8,2)
barplot(porc_p8, porc_p8, col=2:9)
legend(x ="topright", legend = names(p8), fill = 2:9)
```


